---
title: "Cell sampling and Shannon entropy for the 230316 dataset, m367 vs m369"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: yes
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---



```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
PROJECT <- "230316_Kopplin"
outdir <- "/media/hieunguyen/HNSD_MBPro/CRC1382/outdir/LKopplin_OFFICIAL"

path.to.main.src.project <- "/home/hieunguyen/CRC1382/src_2023/LKopplin/230316_Kopplin"
source(file.path(path.to.main.src.project, "00_import_libraries.R"))
source(file.path(path.to.main.src.project, "00_helper_functions.R"))

path.to.main.input <- file.path(outdir, sprintf("%s/1st_round/pct_mito_10_1", PROJECT))
path.to.main.output <- file.path(outdir, sprintf("%s/1st_round/pct_mito_10_1/data_analysis", PROJECT))

path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.02.output <- file.path(path.to.main.output, "02_output_m367_vs_m369")
path.to.03.output <- file.path(path.to.main.output, "03_output_m367_vs_m369")
path.to.04.output <- file.path(path.to.main.output, "04_output_m367_vs_m369")

s.obj <- readRDS(file.path(path.to.02.output, "230316_Kopplin.m367_m369.seurat.obj.addedShannonEntropy.rds"))

path.to.04.output <- file.path(path.to.main.output, "04_output")
dir.create(path.to.04.output, showWarnings = FALSE, recursive = TRUE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
##### read file clone.summary in
clone.summary <- read.csv(file.path(path.to.02.output, "final_VDJ_summary.csv"))
colnames(clone.summary) <- c("X", "CTaa", "m367", "m369", "total", "status", "Shannon.entropy")

##### add total number of cells information to the main seurat object
meta.data <- s.obj@meta.data %>% rownames_to_column("barcode")

meta.data <- merge(meta.data, clone.summary, by.x = "CTaa", by.y = "CTaa", all.x = TRUE) %>%
  column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj@meta.data), ]

s.obj <- AddMetaData(object = s.obj, metadata = meta.data$total, col.name = "total")

###### select clones with >= 10 cells
clone.summary.lt.10 <- subset(clone.summary, clone.summary$total < 10)
clone.summary <- subset(clone.summary, clone.summary$total >= 10)
```

# Sampling cells and calculate Shannon entropy

## Boxplot: All Shannon entropies

```{r dataset_230316_m367_m369_all_read_clone_shannon, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
all.shannon.entropy <- subset(clone.summary, select = c(CTaa, Shannon.entropy))
all.shannon.entropy %>% ggplot(aes(y = Shannon.entropy)) + geom_boxplot()

```

## Boxplot: Comparson between real clones and sampled clones Shannon entropies {.tabset}

```{r dataset_230316_m367_m369_sampling_shannon, ampling_shannon_entropies, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=10, dpi=300, dev='svg'}
all.clone.sizes <- sort(unique(clone.summary$total))
dir.create(file.path(path.to.04.output, "validation_m367_vs_m369"), showWarnings = FALSE, recursive = TRUE)
##### Helper function
calculate_shannon_entropy_sampling_cells <- function(list_of_cells, s.obj){
  N <- length(unique(s.obj$seurat_clusters))
  count_clone_in_cluster <- table(subset(s.obj@meta.data, row.names(s.obj@meta.data) %in% list_of_cells) %>%
                                    subset(select = c(seurat_clusters))) %>%  as.data.frame()
  count_clone_in_cluster <- count_clone_in_cluster %>% rowwise %>% mutate(Prob = Freq / sum(count_clone_in_cluster$Freq)) %>%
    subset(Prob != 0)
  shannon_entropy <- -sum(count_clone_in_cluster$Prob * log2(count_clone_in_cluster$Prob))/log2(N)
  return(shannon_entropy)
}
for (num.sampling.cells in all.clone.sizes){
  all.cells <- colnames(s.obj)
  all.sampling.entropies <- c()

  for (count in seq(1, 1000)){
    sampling.cells <- sample(all.cells, num.sampling.cells)
    tmp.sampling.shannon.entropy <- calculate_shannon_entropy_sampling_cells(sampling.cells, s.obj)
    all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
  }

  samplingdf <- data.frame(data = all.sampling.entropies)
  colnames(samplingdf) <- c("Shannon.entropy")
  samplingdf$CTaa <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
  samplingdf <- samplingdf[c("CTaa", "Shannon.entropy")]
  samplingdf$case <- "Sampling"

  all.shannon.entropy$case <- "real_clone"

  validationdf <- rbind(all.shannon.entropy, samplingdf)

  p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
    ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells))
  cat(sprintf("### Sampling %s cells \n", num.sampling.cells))
  print(p)
  cat("\n \n")
  writexl::write_xlsx(validationdf, file.path(path.to.04.output, "validation_m367_vs_m369", sprintf("validation_Shannon_entropy_sampling_%s.4th_m367_vs_m369_Dataset.xlsx", num.sampling.cells)))
}
```


# Double y-axis plot

- y-axis 1: Shannon entropy.

- y-axis 2: Number of cells in a clone.

## All clones
```{r dataset_230316_m367_m369_double_y_all_clones, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 1000
p <- clone.summary %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) +
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +  
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p
```

## Shared clones between 2 samples
```{r dataset_230316_m367_m369_double_y_clone_in_all_2_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, , dpi=300, dev='svg'}
coeff <- 1000
p <- subset(clone.summary, clone.summary$status == "shared_in_2_samples") %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p
```

## Unique clones
```{r dataset_230316_m367_m369_double_y_clone_in_unique_1_sample, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, , dpi=300, dev='svg'}
coeff <- 1000
p <- subset(clone.summary, clone.summary$status == "unique_in_1_sample") %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) +
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +  
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p

```

# Clone-size - Shannon entropy plot

## UMAP 

### All cells
```{r dataset_230316_m367_m369_UMAP_all_2_samples, echo=FALSE, warning=FALSE, message=FALSE, include=TRUE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12)
```

# New Shannon entropy plot

## Clones >= 10 cells, Shannon entropy on UMAP

Clones < 10 cells are displayed in light grey color. 

```{r dataset_230316_m367_m369_shannon, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", features = c("Shannon.entropy"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#FF0000", na.value = "lightgray") + ggtitle("")

```


## Clones < 10 cells, Clone size on UMAP
Display clone size of all clones < 10 cells on UMAP. 

```{r dataset_230316_m367_m369_clonesize_lt10, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.lt10 = ifelse(CTaa %in% clone.summary.lt.10$CTaa == TRUE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj@meta.data), ]

s.obj <- AddMetaData(object = s.obj, metadata = meta.data$total.lt10, col.name = "total.lt10")
FeaturePlot(object = s.obj, 
            reduction = "INTE_UMAP", features = c("total.lt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

## Clones >= 10 cells, Clone size on UMAP
Display clone size of all clones >= 10 cells on UMAP. 

```{r dataset_230316_m367_m369_clonesize_gt10, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.gt10 = ifelse(CTaa %in% clone.summary.lt.10$CTaa == FALSE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj@meta.data), ]

s.obj <- AddMetaData(object = s.obj, metadata = meta.data$total.gt10, col.name = "total.gt10")
FeaturePlot(object = s.obj, 
            reduction = "INTE_UMAP", features = c("total.gt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

## All cells, Clone size on UMAP

```{r dataset_230316_m367_m369_clonesize_all, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
FeaturePlot(object = s.obj, 
            reduction = "INTE_UMAP", features = c("total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

## All cells, Clone size on UMAP, log

```{r dataset_230316_m367_m369_clonesize_all_logScale, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
s.obj$log.total <- log10(s.obj$total)
FeaturePlot(object = s.obj, 
            reduction = "INTE_UMAP", features = c("log.total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14}
```


