---
title: "Downstream analysis 230215_Kopplin_Pabst, after removing cluster 9 (contaminated cells)"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: true
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    theme: lumen
params:
  PROJECT: NA
  outdir: NA
---
  
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {
  $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
  // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.height=10, fig.width=14}
##### clean up #####
# gc()
# rm(list = ls())
PROJECT <- "230215_Kopplin_Pabst_added_NC_000001_merged_zcat_m330_m331_remove_c9"

path.to.main.src.project <- "/home/hieunguyen/CRC1382/src_2023/LKopplin/230215_Kopplin_Pabst"
source(file.path(path.to.main.src.project, "00_import_libraries.R"))
source(file.path(path.to.main.src.project, "00_helper_functions.R"))

outdir <- "/media/hieunguyen/HNSD_MBPro/CRC1382/outdir/LKopplin_OFFICIAL"
path.to.main.input <- file.path(outdir, sprintf("%s/1st_round/pct_mito_10_1", PROJECT))
path.to.main.output <- file.path(outdir, sprintf("%s/1st_round/pct_mito_10_1/data_analysis", PROJECT))

path.to.02.output <- file.path(path.to.main.output, "02_output")
dir.create(file.path(path.to.02.output), showWarnings = FALSE, recursive = TRUE)

s.obj <- readRDS(file.path(path.to.main.input, "s8_output", sprintf("%s.output.s8.rds", PROJECT)))

##### generate RNA UMAP for each sample #####
if (file.exists(file.path(path.to.02.output, "s.obj.before.integration.rds")) == FALSE){
  s.obj.before.integration <- readRDS(file.path(path.to.main.input, "s6_output", sprintf("%s.output.s6.rds", PROJECT)))
  
  ##### Run s8a source
  chosen.assay <- "RNA"
  DefaultAssay(s.obj.before.integration) <- chosen.assay
  num.PCA <- 25
  num.PC.used.in.UMAP <- 25
  num.PC.used.in.Clustering <- 25
  num.dim.integration <- 25
  num.dim.cluster <- 25
  my_random_seed <- 42
  cluster.resolution <- 0.5
  
  s.obj.before.integration <- RunPCA(s.obj.before.integration, npcs = num.PCA, verbose = FALSE, reduction.name=sprintf("%s_PCA", chosen.assay))
    
  s.obj.before.integration <- RunUMAP(s.obj.before.integration, reduction = sprintf("%s_PCA", chosen.assay), 
                     dims = 1:num.PC.used.in.UMAP, reduction.name=sprintf("%s_UMAP", chosen.assay),
                     seed.use = my_random_seed, umap.method = "uwot")
    
  # clustering 
  s.obj.before.integration <- FindNeighbors(s.obj.before.integration, reduction = sprintf("%s_PCA", chosen.assay), dims = 1:num.PC.used.in.Clustering)
    
  s.obj.before.integration <- FindClusters(s.obj.before.integration, 
                          resolution = cluster.resolution, random.seed = 0)
  
  saveRDS(object = s.obj.before.integration, file = file.path(path.to.02.output, "s.obj.before.integration.rds"))
} else {
  s.obj.before.integration <- readRDS(file.path(path.to.02.output, "s.obj.before.integration.rds"))
}

```

# Preprocessing and QC for GEX data
## Raw data Quality control  {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
for (plot.name in names(s.obj@misc$all.QC)){
 
    cat('###',plot.name,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
    print(s.obj@misc$all.QC[plot.name])
    cat(' \n \n')
}
```



## Ambient RNA background correction 

### DecontX clusters {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
for (plot.name in names(s.obj@misc$ambient.cluster.RNA.plot)){
 
    cat('####',plot.name,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
    print(s.obj@misc$ambient.cluster.RNA.plot[plot.name])
    cat(' \n \n')
}
```

### Contamination level in each sample

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
ggplot(s.obj@meta.data, aes(x=name, y=AmbientRNA)) + 
  geom_boxplot()

```


### Ambient RNA contamination level {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
for (plot.name in names(s.obj@misc$ambient.contamination.plot)){
 
    cat('####',plot.name,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
    print(s.obj@misc$ambient.contamination.plot[plot.name])
    cat(' \n \n')
}
```


## Descriptive statistics and filtering threshold {.tabset}
This section is devoted to the descriptive statistics of the following varialbes: `nFeature_RNA, nCount_RNA, percent.mt, percent.ribo`. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
for (plot.item in c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo")){
  boxplot <- s.obj@meta.data %>% ggplot(aes_string(x = "name", y = plot.item)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16,
               outlier.size=2, notch=FALSE) +
    ggtitle(sprintf("Boxplot: Distribution of %s in each dataset", plot.item))
  cat('###', plot.item,'{.unlisted .unnumbered} \n')
    
    ##### 
    # plots or tables that we want to show in tabs
    #####
  
  print(boxplot)
  
  cat(' \n \n')
}
```



## Descriptive statistics + UMAP {.tabset}

### % Mitochondrial
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "percent.mt", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % Ribosome
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "percent.ribo", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % nCount RNA
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "nCount_RNA", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % nGenes 
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
FeaturePlot(object = s.obj, reduction = "INTE_UMAP", feature = "nFeature_RNA", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

## Cell cycle scoring {.tabset}

### Cell cycle, split by Phase
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
DimPlot(s.obj, reduction = "INTE_UMAP", split.by = "Phase")
```

### Cell cycle, group by Phase
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
DimPlot(s.obj, reduction = "INTE_UMAP", group.by = "Phase", label = TRUE, label.size = 8, pt.size = 0.5, label.box = TRUE, repel = TRUE)
```

### PCA, cell cycle, group by Phase
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
all.genes <- rownames(x = s.obj)
s.genes <- paste0("^", cc.genes$s.genes, "$", collapse = "|")
s.genes <- all.genes[grepl(s.genes, all.genes, ignore.case = TRUE)]
g2m.genes <- paste0("^", cc.genes$g2m.genes, "$", collapse = "|")
g2m.genes <- all.genes[grepl(g2m.genes, all.genes, ignore.case = TRUE)]
s.obj <- RunPCA(s.obj, features = c(s.genes, g2m.genes), nfeatures.print = 10, reduction.name="CELLCYCLED_PCA")

DimPlot(s.obj, reduction = "CELLCYCLED_PCA", group.by = "Phase", pt.size = 1)
```

# Dimension reduction with UMAP

## UMAP: all clusters 
```{r UMAP_all_clusters_after_integrating, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
# DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE) + 
#   ggtitle(sprintf("UMAP: All clusters")) +
#   theme(axis.text = element_text(size=20, face = "bold"),
#         axis.title = element_text(size=20, face = "bold"), 
#         title = element_text(size = 20, face = "bold"))
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE) + ggtitle("")
```

## UMAP: all samples before integrating
```{r UMAP_all_samples_before_integrating, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
# DimPlot(object = s.obj.before.integration, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE, group.by = "name") + 
#   ggtitle(sprintf("UMAP Sample: All samples before integrating")) +
#   theme(axis.text = element_text(size=20, face = "bold"),
#         axis.title = element_text(size=20, face = "bold"), 
#         title = element_text(size = 20, face = "bold"))

DimPlot(object = s.obj.before.integration, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE, group.by = "name") + ggtitle("")
```

## UMAP: all samples after integrating
```{r UMAP_all_samples_after_integrating, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
# DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE, group.by = "name") + 
#   ggtitle(sprintf("UMAP Sample: All samples after integrating")) +
#   theme(axis.text = element_text(size=20, face = "bold"),
#         axis.title = element_text(size=20, face = "bold"), 
#         title = element_text(size = 20, face = "bold"))

DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE, group.by = "name") + ggtitle("")
```

## Distribution of cells having "GRCh38_CD2" gene (or added sequences)

### Where are GRCh38_CD2 (or added sequences) cells?
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}

if (PROJECT %in% c("230215_Kopplin_Pabst_NC_000001", "230215_Kopplin_Pabst_NC_000001_removed_contaminated_cells",
                         "230215_Kopplin_Pabst_added_NC_000001_merged", "230215_Kopplin_Pabst_added_NC_000001_merged_zcat",
                         "230215_Kopplin_Pabst_added_NC_000001_merged_zcat_m330_m331",
                         "230215_Kopplin_Pabst_added_NC_000001_merged_zcat_m330_m331_remove_c9")){
  added.gene <- "NC-000001.11"
} else if (PROJECT %in% c("230215_Kopplin_Pabst_NC_060925", "230215_Kopplin_Pabst_added_NC_060925_merged", 
                                "230215_Kopplin_Pabst_added_NC_060925_merged_zcat")){
  added.gene <- "NC-060925.1"
} else if (PROJECT == "230215_Kopplin_Pabst"){
  added.gene <- "GRCh38-CD2"
} else if (PROJECT == "230215_Kopplin_Pabst_removed_contaminated_cells"){
  added.gene <- "GRCh38-CD2"
}

count.mat.CD2 <- GetAssayData(object = s.obj, slot = "counts", assay = "RNA")[added.gene, ]
cd2.cells <- names(count.mat.CD2[count.mat.CD2!= 0])

DimPlot(object = s.obj, reduction = "INTE_UMAP", cells.highlight = cd2.cells, cols.highlight = c("red"), label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 0.5, sizes.highlight = 1) + theme(legend.position = "none") + 
  ggtitle(sprintf("Distribution of cells having %s gene, # = %s", added.gene, length(cd2.cells)))
```

## Expressions of GRCh38_CD2 in each cell
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
DefaultAssay(s.obj) <- "RNA"
FeaturePlot(object = s.obj, features = c(added.gene), label = TRUE, pt.size = 1)
```


# Cluster marker genes
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
if (file.exists(file.path(path.to.02.output, "all_cluster_markers.rds")) == FALSE){
  cluster.markers <- FindAllMarkers(object = s.obj, assay = "RNA", test.use = "wilcox")
  cluster.markers <- subset(cluster.markers, cluster.markers$p_val_adj <= 0.05 & cluster.markers$avg_log2FC > 0)
  saveRDS(cluster.markers, file.path(path.to.02.output, "all_cluster_markers.rds"))
} else {
  cluster.markers <- readRDS(file.path(path.to.02.output, "all_cluster_markers.rds"))
}

```

## Tables: All cluster marker genes {.tabset}

```{r echo=FALSE, warning=FALSE, results='asis', include=FALSE}
cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dev='svg'}
for(cluster.id in sort(unique(cluster.markers$cluster))){
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  tmp <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC)) %>%
    mutate_if(is.numeric, round, 6)
  row.names(tmp) <- NULL
  print( htmltools::tagList(datatable(tmp, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}
```

## Top-16 cluster markers feature plot {.tabset}

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10, dev='svg'}
for (cluster.id in unique(cluster.markers$cluster)){
  DefaultAssay(s.obj) <- "RNA"
  tmp <- subset(cluster.markers, cluster.markers$cluster == cluster.id & cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>% head(16)
  cat(sprintf("### Cluster %s \n", cluster.id))
  p <- FeaturePlot(object = s.obj, reduction = "INTE_UMAP", features = tmp$gene, ncol = 4, ) 
  print(p)
  cat("\n \n")
}
```

## Top-16 cluster markers dot plot {.tabset}

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10, dev='svg'}
for (cluster.id in unique(cluster.markers$cluster)){
  DefaultAssay(s.obj) <- "RNA"
  tmp <- subset(cluster.markers, cluster.markers$cluster == cluster.id & cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>% head(16)
  cat(sprintf("### Cluster %s \n", cluster.id))
  p <- DotPlot(object = s.obj, features = tmp$gene) 
  print(p)
  cat("\n \n")
}
```

## Top-16 cluster markers Violin plot {.tabset}

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10, dev='svg'}
for (cluster.id in unique(cluster.markers$cluster)){
  DefaultAssay(s.obj) <- "RNA"
  tmp <- subset(cluster.markers, cluster.markers$cluster == cluster.id & cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>% head(16)
  cat(sprintf("### Cluster %s \n", cluster.id))
  p <- VlnPlot(object = s.obj, features = tmp$gene, pt.size = 0) 
  print(p)
  cat("\n \n")
}
```

# VDJ data

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
path.to.VDJ.output <- file.path(outdir, PROJECT, "VDJ_output")

all.vdj.files <- Sys.glob(file.path(path.to.VDJ.output, "annotated_contigs_clonaltype_*.csv"))
all.vdj.info.df <- data.frame()
for (file in all.vdj.files){
  tmpdf <- read.csv(file) %>% subset(select = -c(X)) 
  all.vdj.info.df <- rbind(all.vdj.info.df, tmpdf)
}
```

## Full VDJ data tables {.tabset}
```{r echo=FALSE, warning=FALSE, results='asis', include=FALSE}
all.vdj.info.df %>% create_dt()

all.vdj <- hash()
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
for (sample.id in unique(all.vdj.info.df$sample)){
  tmpdf <- subset(all.vdj.info.df, all.vdj.info.df$sample == sample.id)
  all.vdj[[sample.id]] <- tmpdf
  cat(paste("\n\n### sample: ", sample.id, "##\n"))
  print( htmltools::tagList(datatable(tmpdf, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}
```


## Number of cells in each clone in each sample {.tabset}
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=12}
#### merge VDJ information to the main Seurat object
meta.data <- s.obj@meta.data %>% rownames_to_column("barcode")
meta.data <- merge(meta.data, all.vdj.info.df, by.x = "barcode", by.y = "barcode", all.x = TRUE)
s.obj <- AddMetaData(object = s.obj, metadata = meta.data$CTaa, col.name = "CTaa")

subset.s.obj <- hash()
for (sample.id in unique(s.obj$name)){
  cat(sprintf("### Sample: %s \n", sample.id))
  subset.s.obj[[sample.id]] <- subset(s.obj, name == sample.id)  
  cells.with.CTaa <- row.names(subset(subset.s.obj[[sample.id]]@meta.data, is.na(subset.s.obj[[sample.id]]@meta.data$CTaa) == FALSE))
  p <- DimPlot(object = subset.s.obj[[sample.id]], reduction = "INTE_UMAP", cells.highlight = cells.with.CTaa, pt.size = 0.8, cols.highlight = "red") + ggtitle(sprintf("Sample: %s, # cells with clone: %s, Total # cells: %s , percentage = %s:", sample.id, length(cells.with.CTaa), length(colnames(subset.s.obj[[sample.id]])), round(100*length(cells.with.CTaa)/length(colnames(subset.s.obj[[sample.id]])), 3))) + theme(legend.position = "none")
  print(p)
  cat("\n \n")
}


```

## Number of appearances of each clone in each sample
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
all.clone.df <- data.frame(unique(all.vdj.info.df$CTaa))
colnames(all.clone.df) <- c("clone")
all.clone.df <- subset(all.clone.df, is.na(all.clone.df$clone) == FALSE)
all.clone.df <- all.clone.df %>% rowwise %>% 
  mutate(m330_merged = nrow(subset(subset.s.obj[["m330_merged"]]@meta.data, subset.s.obj[["m330_merged"]]@meta.data$CTaa == clone))) %>%
  mutate(m331_merged = nrow(subset(subset.s.obj[["m331_merged"]]@meta.data, subset.s.obj[["m331_merged"]]@meta.data$CTaa == clone))) %>%
  mutate(total = m330_merged + m331_merged)
all.clone.df %>% create_dt()
```

## Unique clones

Clones with total occurrences >= 5

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
unique.clones <- hash()

unique.clones[["m330_merged"]] <- subset(all.clone.df, all.clone.df$m330_merged != 0 & all.clone.df$m331_merged == 0)
unique.clones[["m331_merged"]] <- subset(all.clone.df, all.clone.df$m330_merged == 0 & all.clone.df$m331_merged != 0)
```


```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=12}
for (sample.id in names(unique.clones)){
  cat(sprintf("### Sample %s {.tabset} \n", sample.id))
  
  list.of.clones <- unique.clones[[sample.id]] %>% subset(total >= 5)
  split.list.of.clones <- split(list.of.clones$clone, ceiling(seq_along(list.of.clones$clone) / 6))
  for (group in names(split.list.of.clones)){
    cat(sprintf("#### Part %s \n", group))
    p <- plot_a_list_of_clonal_types(s.obj = s.obj, list_of_clonal_types = split.list.of.clones[[group]])
    print(p)
    cat("\n \n")
  }
}

```

## Clones shared in 2 samples
Clones with total occurrences >= 5

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
shared2.clones <- hash()
shared2.clones[["m330_merged_m331_merged"]] <- subset(all.clone.df, all.clone.df$m330_merged != 0 & all.clone.df$m331_merged != 0 )
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=12}
for (sample.id in names(shared2.clones)){
  cat(sprintf("### Sample %s {.tabset} \n", sample.id))
  
  list.of.clones <- shared2.clones[[sample.id]] %>% subset(total >= 5)
  split.list.of.clones <- split(list.of.clones$clone, ceiling(seq_along(list.of.clones$clone) / 6))
  for (group in names(split.list.of.clones)){
    cat(sprintf("#### Part %s \n", group))
    p <- plot_a_list_of_clonal_types(s.obj = s.obj, list_of_clonal_types = split.list.of.clones[[group]])
    print(p)
    cat("\n \n")
  }
}

```

## Shannon entropy calculation (cluster-based)
```{r echo=FALSE, warning=FALSE, message=FALSE}
#####----------------------------------------------------------------------#####
# Calculate SHANNON ENTROPY
#####----------------------------------------------------------------------#####

##### higher Shannon entropy --> more diverse
calculate_shannon_entropy <- function(clone, s.obj){
  N <- length(unique(s.obj$seurat_clusters))
  count_clone_in_cluster <- table(subset(s.obj@meta.data, s.obj@meta.data$CTaa == clone) %>% 
                                    subset(select = c(CTaa, seurat_clusters))) %>%  as.data.frame()
  count_clone_in_cluster <- count_clone_in_cluster %>% rowwise %>% mutate(Prob = Freq / sum(count_clone_in_cluster$Freq)) %>%
    subset(Prob != 0)
  shannon_entropy <- -sum(count_clone_in_cluster$Prob * log2(count_clone_in_cluster$Prob))/log2(N)
  return(shannon_entropy)
}

final.VDJ.summary <- data.frame()

for (group in names(unique.clones)){
  tmp <- unique.clones[[group]]
  tmp[["status"]] <- "unique_in_1_sample"
  final.VDJ.summary <- rbind(final.VDJ.summary, tmp)
}

for (group in names(shared2.clones)){
  tmp <- shared2.clones[[group]]
  tmp[["status"]] <- "shared_in_2_samples"
  final.VDJ.summary <- rbind(final.VDJ.summary, tmp)
}

final.VDJ.summary <- as.data.frame(final.VDJ.summary)

shannon.entropies <- unlist(lapply(final.VDJ.summary$clone, function(x){
  return(calculate_shannon_entropy(clone = x, s.obj = s.obj))
}))

final.VDJ.summary[["Shannon.entropy"]] <- shannon.entropies
```

### Summary table 
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
final.VDJ.summary  %>% arrange(desc(total)) %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=15, fig.height=10}
shannon.entropy.df <- final.VDJ.summary %>% subset(select = c(clone, Shannon.entropy))

meta.data <- s.obj@meta.data

cell.shannon.entropy <- unlist(lapply(meta.data$CTaa, function(x){
  if (is.na(x) == FALSE){
    return(as.numeric(subset(shannon.entropy.df, shannon.entropy.df$clone == x)$Shannon.entropy))  
  } else {
    return(NA)
  }
}))
s.obj <- AddMetaData(object = s.obj, metadata = cell.shannon.entropy, col.name= "Shannon.entropy")

FeaturePlot(subset(s.obj,Shannon.entropy != "NA"), reduction = "INTE_UMAP", features = c("Shannon.entropy"))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
##### SAVE OBJECTS 
saveRDS(object = s.obj, file.path(path.to.02.output, "230215_Kopplin.seurat.obj.addedShannonEntropy.rds"))
write.csv(final.VDJ.summary, file.path(path.to.02.output, "final_VDJ_summary.csv"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
```

