---
title: "GEX data analysis for the second dataset"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---



```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
#####----------------------------------------------------------------------#####
#
# 06: Increase higher resolution for the clustering of the first dataset
#
# trnguyen@ukaachen.de
#####----------------------------------------------------------------------#####

##### clean up #####
# gc()
# rm(list = ls())

path.to.pipeline.src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline"
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "import_libraries.R"))
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "helper_functions.R"))

path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/LKopplin/20231101_OFFICIAL/Dataset2"
source(file.path(path.to.project.src, "helper_functions.R"))

#####----------------------------------------------------------------------#####
# CONFIGURATIONS 
#####----------------------------------------------------------------------#####

chosen.seed <- 42
num.dim.integration <- 25 
num.PCA <- 25
num.dim.cluster <- 25
num.PC.used.in.Clustering <- 25

source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "s8_integration_and_clustering.R"))

PROJECT <- "211227_Kopplin"
outdir <- "/media/hieunguyen/HNSD_MBPro/CRC1382/outdir/LKopplin_OFFICIAL"

path.to.main.output <- file.path(outdir, PROJECT, "data_analysis")
path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.02.output <- file.path(path.to.main.output, "02_output")
dir.create(path.to.02.output, showWarnings = FALSE, recursive = TRUE)

s.obj <- readRDS(file.path(path.to.01.output, "merged_all_second_exp_dataset.annotated.filteredCD45.integrated.rds"))

if ("svglite" %in% installed.packages() == FALSE){
  install.packages("svglite")
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
label <- TRUE
label.box <- TRUE
label.color <- "blue"
label.size <- 8
repel <- TRUE 
pt.size <- 0.5
fig.width <- 14
fig.height <- 10
cols <- c("lightgray", "darkblue")

print(sprintf("Label on UMAP: %s", label))
print(sprintf("Label box on UMAP: %s", label))
print(sprintf("Label color: %s", label.color))
print(sprintf("Label size: %s", label.size))
print(sprintf("Point size: %s", pt.size))
print(sprintf("Fig width x height: %s x %s", fig.width, fig.height))
print(sprintf("Color bar: %s - %s", cols[[1]], cols[[2]]))
```

## UMAP, group by current clusters
```{r UMAP_integration_2nd_dataset_groupby_clusters, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dpi = 300, dev = 'svg'}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = label, label.box = TRUE, repel = TRUE, pt.size = pt.size) +
  theme(text = element_text(size=25),
        axis.text.x = element_text(hjust=1)) + 
  xlim(-7, 10) + ylim(-8, 8)
```

## UMAP, group by samples
```{r UMAP_integration_2nd_dataset_groupby_sample, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dpi = 300, dev = 'svg'}
DimPlot(object = s.obj, group.by = "name", reduction = "INTE_UMAP", label = label, label.box = TRUE, repel = TRUE, pt.size = pt.size) + theme(text = element_text(size=25),
        axis.text.x = element_text(hjust=1)) + 
  xlim(-7, 10) + ylim(-8, 8)
```


# Cluster markers
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10, dpi = 300, dev = 'svg'}
# Run find all cluster markers with default input parameters
if (file.exists(file.path(path.to.02.output, "cluster_markers.rds")) == FALSE){
  cluster.markers <- FindAllMarkers(object = s.obj, test.use = "wilcox", assay = "RNA")
  cluster.markers <- subset(cluster.markers, (cluster.markers$p_val_adj <= 0.05) & (cluster.markers$avg_log2FC > 0))
  saveRDS(cluster.markers, file.path(path.to.02.output, "cluster_markers.rds"))  
} else {
  cluster.markers <- readRDS(file.path(path.to.02.output, "cluster_markers.rds"))
}
DefaultAssay(s.obj) <- "RNA"
```

### Tables of all cluster marker genes {.tabset}
```{r echo=FALSE, include=FALSE}
cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=30, fig.height=20}

for (cluster.id in sort(unique(cluster.markers$cluster))){
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  tmp.cluster.markers <- tmp.cluster.markers %>% mutate_if(is.numeric, round, 6) 
  
  cat(paste("\n\n#### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.cluster.markers, extensions = 'Buttons',
                                      filter = "top",
                                      options = list(dom = 'Blfrtip',
                                                     buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                     lengthMenu = list(c(10,25,50,-1),
                                                                       c(10,25,50,"All")),
                                                     columnDefs = list(list(
                                                       targets = "_all",
                                                       render = JS(
                                                         "function(data, type, row, meta) {",
                                                         "return type === 'display' && data != null && data.length > 100 ?",
                                                         "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                                         "}")
                                                     ))
                                      ))))
  cat("\n \n")
}

```


### Top-12 cluster marker genes `FeaturePlot` {.tabset}
Here we just show the top-12 genes (ranked by `avg_log2FC`) for each cluster. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=15}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("#### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- FeaturePlot(object = s.obj, reduction = "RNA_UMAP", features = head(tmp.cluster.markers, 12)$gene, ncol = 3, label = TRUE, pt.size = 0.5, label.size = 5, label.color = "red")  
  print(p)
  cat("\n \n")
}
```

### Top-12 cluster marker genes `DotPlot` {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=15}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("#### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- DotPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

### Top-12 cluster marker genes `Vlnplot` {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=15}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("#### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- VlnPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

# Heatmap: Cluster markers
```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=30, fig.height=18, dpi = 300}
library(viridis)

list.of.top10.cluster.markers <- data.frame()
for (cluster.id in unique(cluster.markers$cluster)){
  tmp <- subset(cluster.markers, (cluster.markers$cluster == cluster.id) & (cluster.markers$p_val_adj <= 0.05) & (cluster.markers$avg_log2FC > 0)) %>% arrange(desc(avg_log2FC)) %>% head(10)
  list.of.top10.cluster.markers <- rbind(list.of.top10.cluster.markers, tmp)
}

gex_cluster_markers_heatmap <- DoHeatmap(object = s.obj, features = list.of.top10.cluster.markers$gene, label = TRUE, draw.lines = T, assay = "RNA") + scale_fill_viridis(direction = -1)

library(svglite)
ggsave(plot = gex_cluster_markers_heatmap, filename = file.path(path.to.02.output, sprintf("heatmap_all_top10_cluster_marker_genes.svg")),
       width = 30, height = 18, dpi = 400, device = 'png')

print(gex_cluster_markers_heatmap)
```

# Clone VDJ information in all three samples 
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=14, fig.width=10}
#####----------------------------------------------------------------------#####
# PROCESSING CLONES FROM THE SECOND DATASET
#####----------------------------------------------------------------------#####
if (file.exists(file.path(path.to.02.output, "final_summary_clones_2ndDataset.xlsx")) == FALSE){
  count.clonaltype.2nd <- hash()
  
  for (sample.id in unique(s.obj$name)){
    tmpdf <- table(subset(s.obj, name == sample.id)$CTaa) %>% as.data.frame()
    colnames(tmpdf) <- c("CTaa", "count")
    count.clonaltype.2nd[[sample.id]] <- tmpdf
  }
  
  # count the appearance of all clonal types in all samples
  tmp1 <- merge(count.clonaltype.2nd$GFP_m1, count.clonaltype.2nd$GFP_m2, 
                by.x = "CTaa", by.y = "CTaa", 
                all.x = TRUE, all.y = TRUE) %>% arrange(desc(count.x))
  
  summary.count.clonaltype.2nd <- merge(tmp1, count.clonaltype.2nd$GFP_m3, 
                                    by.x = "CTaa", by.y = "CTaa", 
                                    all.x = TRUE, all.y = TRUE) %>% arrange(desc(count.x))
  
  colnames(summary.count.clonaltype.2nd) <- c("CTaa", "GFP_m1", "GFP_m2", "GFP_m3")
  summary.count.clonaltype.2nd[is.na(summary.count.clonaltype.2nd)] <- 0
  
  summary.clonaltype.2nd <- data.frame(unique(c(count.clonaltype.2nd$GFP_m1$CTaa, 
                                            count.clonaltype.2nd$GFP_m2$CTaa, 
                                            count.clonaltype.2nd$GFP_m3$CTaa)))
  
  colnames(summary.clonaltype.2nd) <- c("CTaa")
  
  summary.clonaltype.2nd <- summary.clonaltype.2nd %>% rowwise %>% 
    mutate(Status = assign_status_for_clonaltype_2nd(clone = CTaa, 
                                                     count.clonaltype = count.clonaltype.2nd, 
                                                     summary.count.clonaltype = summary.count.clonaltype.2nd))
  
  final.summary.clonaltype.2nd <- merge(summary.clonaltype.2nd, summary.count.clonaltype.2nd, by.x = "CTaa", by.y = "CTaa")
  final.summary.clonaltype.2nd.raw <- final.summary.clonaltype.2nd
  final.summary.clonaltype.2nd <- subset(final.summary.clonaltype.2nd, final.summary.clonaltype.2nd$Status != "excluded")
  
  final.summary.clonaltype.2nd <- final.summary.clonaltype.2nd %>% rowwise %>%
    mutate(total = GFP_m1 + GFP_m2 + GFP_m3) 
  
  writexl::write_xlsx(x = final.summary.clonaltype.2nd.raw, path = file.path(path.to.02.output, "final_summary_clones_2ndDataset.raw.xlsx"))
  writexl::write_xlsx(x = final.summary.clonaltype.2nd, path = file.path(path.to.02.output, "final_summary_clones_2ndDataset.xlsx"))
} else {
  final.summary.clonaltype.2nd.raw <- readxl::read_xlsx(file.path(path.to.02.output, "final_summary_clones_2ndDataset.raw.xlsx"))
  final.summary.clonaltype.2nd <- readxl::read_xlsx(file.path(path.to.02.output, "final_summary_clones_2ndDataset.xlsx"))
}

```

## Clone VDJ information with Shannon entropy
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=14, fig.width=10}
if (file.exists(file.path(path.to.02.output, "final_summary_clones_2ndDataset.withShannonEntropy.xlsx")) == FALSE){
  summary.clones.2nd <- final.summary.clonaltype.2nd %>% rowwise %>%
  mutate(Shannon.entropy = calculate_shannon_entropy(clone = CTaa, s.obj = s.obj))
writexl::write_xlsx(summary.clones.2nd, file.path(path.to.02.output, "final_summary_clones_2ndDataset.withShannonEntropy.xlsx"))  
} else {
  summary.clones.2nd <- readxl::read_xlsx(file.path(path.to.02.output, "final_summary_clones_2ndDataset.withShannonEntropy.xlsx"))
}
summary.clones.2nd %>% create_dt()
```

# Generate clone-UMAP

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=14, fig.width=10}
#####
# Path to save .svg files for clone-UMAP plots.
#####
dir.create(file.path(path.to.02.output, "shared_3_samples_clones"), showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(path.to.02.output, "unique_clones"), showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(path.to.02.output, "shared_2_samples_clones"), showWarnings = FALSE, recursive = TRUE)

clone.summary <- final.summary.clonaltype.2nd
unique.clones <- subset(clone.summary, clone.summary$Status == "unique_in_1_sample") %>% arrange(desc(total))
shared2.clones <- subset(clone.summary, clone.summary$Status == "in_2_samples") %>% arrange(desc(total))
shared3.clones <- subset(clone.summary, clone.summary$Status == "in_all_3_samples") %>% arrange(desc(total))

split.unique.clones <- split(unique.clones$CTaa, ceiling(seq_along(unique.clones$CTaa) / 9))
split.shared2.clones <- split(shared2.clones$CTaa, ceiling(seq_along(shared2.clones$CTaa) / 9))
split.shared3.clones <- split(shared3.clones$CTaa, ceiling(seq_along(shared3.clones$CTaa) / 9))

for (part in names(split.unique.clones)){
  p <- plot_a_list_of_clonal_types_2nd_dataset(split.unique.clones[[part]], s.obj)
  ggsave(plot = p, filename = file.path(path.to.02.output, "unique_clones", sprintf("unique_clones_UMAP_part_%s", part)),
         device = 'svg', dpi = 300, width = 14, height = 10, units = "in")
}


for (part in names(split.shared2.clones)){
  p <- plot_a_list_of_clonal_types_2nd_dataset(split.shared2.clones[[part]], s.obj)
  ggsave(plot = p, filename = file.path(path.to.02.output, "shared_2_samples_clones", sprintf("shared2_clones_UMAP_part_%s", part)),
         device = 'svg', dpi = 300, width = 14, height = 10, units = "in")
}


for (part in names(split.shared3.clones)){
  p <- plot_a_list_of_clonal_types_2nd_dataset(split.shared3.clones[[part]], s.obj)
  ggsave(plot = p, filename = file.path(path.to.02.output, "shared_3_samples_clones", sprintf("shared3_clones_UMAP_part_%s", part)),
         device = 'svg', dpi = 300, width = 14, height = 10, units = "in")
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=14, fig.width=10}
```



