---
title: "Cell sampling and Shannon entropy for the 2nd dataset"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: yes
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---



```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
##### PREPARATION
##### clean up #####
# gc()
# rm(list = ls())

library(shiny)

path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/LKopplin/20231101_OFFICIAL/Dataset2"
source(file.path(path.to.project.src, "helper_functions_for_03_analysis.R"))
source(file.path(path.to.project.src, "helper_functions.R"))

path.to.pipeline.src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline"
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "import_libraries.R"))
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "helper_functions.R"))
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "s8_integration_and_clustering.R"))

PROJECT <- "211227_Kopplin"
outdir <- "/media/hieunguyen/HNSD_MBPro/CRC1382/outdir/LKopplin_OFFICIAL"
path.to.main.output <- file.path(outdir, PROJECT, "data_analysis")

path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.02.output <- file.path(path.to.main.output, "02_output")
path.to.03.output <- file.path(path.to.main.output, "03_output")
path.to.04.output <- file.path(path.to.main.output, "04_output")
dir.create(path.to.04.output, showWarnings = FALSE, recursive = TRUE)

s.obj.2nd <- readRDS(file.path(path.to.main.output, "01_output" , "merged_all_second_exp_dataset.annotated.filteredCD45.integrated.rds"))

clone.summary <- readxl::read_excel(file.path(path.to.main.output, "02_output", "final_summary_clones_2ndDataset.withShannonEntropy.xlsx"))

clone.summary.raw <-  readxl::read_excel(file.path(path.to.main.output, "02_output", "final_summary_clones_2ndDataset.raw.xlsx"))
clone.summary.raw <- clone.summary.raw %>% rowwise() %>% mutate(total = GFP_m1 + GFP_m2 + GFP_m3)
```

# UMAP

## All samples
```{r dataset2_UMAP_all_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = s.obj.2nd, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12) +
  ggtitle("") 
```


## GFP_m1

```{r dataset2_UMAP_sample_GFP_m1, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = subset(s.obj.2nd, name == "GFP_m1"), reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12) +
  ggtitle("") 
```

## GFP_m2
```{r dataset2_UMAP_sample_GFP_m2, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = subset(s.obj.2nd, name == "GFP_m2"), reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12) +
  ggtitle("") 
```

## GFP_m3
```{r dataset2_UMAP_sample_GFP_m3, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = subset(s.obj.2nd, name == "GFP_m3"), reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12) +
  ggtitle("")
```


# Sampling cells and calculate Shannon entropy

## Boxplot: All Shannon entropies

```{r dataset2_all_read_clone_shannon, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
all.shannon.entropy <- subset(clone.summary, select = c(CTaa, Shannon.entropy))
all.shannon.entropy %>% ggplot(aes(y = Shannon.entropy)) + geom_boxplot() 

```

## Boxplot: Comparson between real clones and sampled clones Shannon entropies {.tabset}

```{r dataset2_sampling_shannon, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
all.clone.sizes <- sort(unique(clone.summary$total))

dir.create(file.path(path.to.04.output, "validation_Shannon_entropy"), showWarnings = FALSE, recursive = TRUE)

##### Helper function
calculate_shannon_entropy_sampling_cells <- function(list_of_cells, s.obj){
  N <- length(unique(s.obj$seurat_clusters))
  count_clone_in_cluster <- table(subset(s.obj@meta.data, row.names(s.obj@meta.data) %in% list_of_cells) %>% 
                                    subset(select = c(seurat_clusters))) %>%  as.data.frame()
  count_clone_in_cluster <- count_clone_in_cluster %>% rowwise %>% mutate(Prob = Freq / sum(count_clone_in_cluster$Freq)) %>%
    subset(Prob != 0)
  shannon_entropy <- -sum(count_clone_in_cluster$Prob * log2(count_clone_in_cluster$Prob))/log2(N)
  return(shannon_entropy)
}

for (num.sampling.cells in all.clone.sizes){
  if (file.exists(file.path(path.to.04.output, "validation_Shannon_entropy", sprintf("validation_Shannon_entropy_sampling_%s.2ndDataset.xlsx", num.sampling.cells))) == FALSE){
    all.cells <- colnames(s.obj.2nd)
    all.sampling.entropies <- c()
    
    for (count in seq(1, 1000)){
      sampling.cells <- sample(all.cells, num.sampling.cells)
      tmp.sampling.shannon.entropy <- calculate_shannon_entropy_sampling_cells(sampling.cells, s.obj.2nd)
      all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
    }
    
    samplingdf <- data.frame(data = all.sampling.entropies)
    colnames(samplingdf) <- c("Shannon.entropy")
    samplingdf$CTaa <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
    samplingdf <- samplingdf[c("CTaa", "Shannon.entropy")]
    samplingdf$case <- "Sampling"
    
    all.shannon.entropy$case <- "real_clone"
    
    validationdf <- rbind(all.shannon.entropy, samplingdf)
    
    writexl::write_xlsx(validationdf, file.path(path.to.04.output, "validation_Shannon_entropy", sprintf("validation_Shannon_entropy_sampling_%s.2ndDataset.xlsx", num.sampling.cells)))
  } else {
    validationdf <- readxl::read_xlsx(file.path(path.to.04.output, "validation_Shannon_entropy", sprintf("validation_Shannon_entropy_sampling_%s.2ndDataset.xlsx", num.sampling.cells)))
  }
  
  p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
    ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells)) 
  cat(sprintf("### Sampling %s cells \n", num.sampling.cells))
  print(p)
  cat("\n \n")
}
```


# Double y-axis plot

## All clones
```{r table_all_clones, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
clone.summary %>% create_dt()
```


```{r dataset2_double_y_all_clones, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 1000
p <- clone.summary %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p

```

## Clones in all 3 samples
```{r dataset2_double_y_clone_in_all_3_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 1000
p <- subset(clone.summary, clone.summary$Status == "in_all_3_samples") %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p

```

## Clones in 2 samples
```{r dataset2_double_y_clone_in_2_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 1000
p <- subset(clone.summary, clone.summary$Status == "in_2_samples") %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p

```


## Clones unique in 1 sample
```{r dataset2_double_y_clone_in_unique_1_sample, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 1000
p <- subset(clone.summary, clone.summary$Status == "unique_in_1_sample") %>% ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p

```

# New Shannon entropy plot
```{r echo=FALSE, include=FALSE, fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj.2nd@meta.data

get_shannon_entropy <- function(x, clone.summary){
  if (nrow( subset(clone.summary, clone.summary$CTaa == x) ) != 0){
    return(subset( clone.summary, clone.summary$CTaa == x )$Shannon.entropy)
  } else {
    return(NA)
  }
}

colnames(clone.summary.raw) <- c("Clone", "Status", "GFP_m1", "GFP_m2", "GFP_m3", "total.raw" )
meta.data <- meta.data %>% rowwise() %>% mutate(Shannon.entropy = get_shannon_entropy(CTaa, clone.summary)) %>%
  mutate(total = ifelse(is.na(CTaa), NA, subset(clone.summary.raw, clone.summary.raw$Clone == CTaa)$total.raw))
  
s.obj.2nd <- AddMetaData(object = s.obj.2nd, metadata = meta.data$Shannon.entropy, col.name = "Shannon.entropy")
s.obj.2nd <- AddMetaData(object = s.obj.2nd, metadata = meta.data$total, col.name = "total")
```

## Clones >= 10 cells, Shannon entropy on UMAP

Clones < 10 cells are displayed in light grey color. 

```{r dataset2_shannon, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
FeaturePlot(object = s.obj.2nd, reduction = "INTE_UMAP", features = c("Shannon.entropy"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#FF0000", na.value = "lightgray") + ggtitle("")
```


## Clones < 10 cells, Clone size on UMAP
Display clone size of all clones < 10 cells on UMAP. 

```{r dataset2_clonesize_lt10, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj.2nd@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.lt10 = ifelse(CTaa %in% clone.summary$CTaa == FALSE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj.2nd@meta.data), ]

s.obj.2nd <- AddMetaData(object = s.obj.2nd, metadata = meta.data$total.lt10, col.name = "total.lt10")
FeaturePlot(object = s.obj.2nd, 
            reduction = "INTE_UMAP", features = c("total.lt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```


## Clones >= 10 cells, Clone size on UMAP
Display clone size of all clones >= 10 cells on UMAP. 

```{r dataset2_clonesize_gt10, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj.2nd@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.gt10 = ifelse(CTaa %in% clone.summary$CTaa == TRUE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj.2nd@meta.data), ]

s.obj.2nd <- AddMetaData(object = s.obj.2nd, metadata = meta.data$total.gt10, col.name = "total.gt10")
FeaturePlot(object = s.obj.2nd, 
            reduction = "INTE_UMAP", features = c("total.gt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

## All cells, Clone size on UMAP

```{r dataset2_clonesize_all, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
FeaturePlot(object = s.obj.2nd, 
            reduction = "INTE_UMAP", features = c("total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

## All cells, Clone size on UMAP. logScale

```{r dataset2_clonesize_all_logScale, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
s.obj.2nd$log.total <- log10(s.obj.2nd$total)
FeaturePlot(object = s.obj.2nd, 
            reduction = "INTE_UMAP", features = c("log.total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
```


