---
title: "Cell sampling and Shannon entropy for the 1st dataset"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: yes
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---



```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
##### PREPARATION
##### clean up #####
# gc()
# rm(list = ls())

path.to.pipeline.src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline"
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "import_libraries.R"))
source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "helper_functions.R"))

path.to.project.src <- "/home/hieunguyen/CRC1382/src_2023/LKopplin/20231101_OFFICIAL/Dataset1"
source(file.path(path.to.project.src, "helper_functions.R"))

#####----------------------------------------------------------------------#####
# CONFIGURATIONS 
#####----------------------------------------------------------------------#####

chosen.seed <- 42
num.dim.integration <- 25 
num.PCA <- 25
num.dim.cluster <- 25
num.PC.used.in.Clustering <- 25

source(file.path(path.to.pipeline.src, "scRNA_GEX_pipeline", "processes_src", "s8_integration_and_clustering.R"))

PROJECT <- "1stExp_Kopplin"
outdir <- "/media/hieunguyen/HNSD_MBPro/CRC1382/outdir/LKopplin_OFFICIAL"

path.to.main.output <- file.path(outdir, PROJECT, "data_analysis")
path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.02.output <- file.path(path.to.main.output, "02_output")
path.to.03.output <- file.path(path.to.main.output, "03_output")
dir.create(path.to.03.output, showWarnings = FALSE, recursive = TRUE)

path.to.save.final.clonedf <- file.path(outdir, "summary_clone_dfs")
dir.create(path.to.save.final.clonedf, showWarnings = FALSE, recursive = TRUE)

cluster.resolution <- 1
s.obj <- readRDS(file.path(path.to.02.output, sprintf("sobj_with_clusterRes_%s.rds", cluster.resolution)))


if ("svglite" %in% installed.packages() == FALSE){
  install.packages("svglite")
}

s.obj.1st.GFP <- subset(s.obj, seurat_clusters %in% c(3, 5, 6, 8))
s.obj.1st.CD45 <- subset(s.obj, seurat_clusters %in% c(3, 5, 6, 8) == FALSE)

dir.create(file.path(path.to.03.output, "validation_CD45_ShannonEntropy"), showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(path.to.03.output, "validation_GFP_ShannonEntropy"), showWarnings = FALSE, recursive = TRUE)
```

# UMAP

## GFP samples

```{r dataset1_UMAP_GFP_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = s.obj.1st.GFP, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12) +
  ggtitle("") 
```

## CD45 samples
```{r dataset1_UMAP_CD45_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
DimPlot(object = s.obj.1st.CD45, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1, label.size = 12) +
  ggtitle("") 
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
##### summary clone information
clone.summary.gfp.modified <- readxl::read_excel(file.path(path.to.02.output, "final_summary_clones_1stDataset_GFP_clusterRes_1.modified.xlsx"))
clone.summary.gfp.modified.raw <- readxl::read_excel(file.path(path.to.02.output, "final_summary_clones_1stDataset_GFP.raw.xlsx")) 
clone.summary.gfp.modified.raw <- clone.summary.gfp.modified.raw %>%
  rowwise() %>%
  mutate(total = GFPexp1_m1 + GFPexp1_m2)
clone.summary.gfp <- readxl::read_excel(file.path(path.to.02.output, "final_summary_clones_1stDataset_GFP.xlsx"))

clone.summary.cd45.modified <- readxl::read_excel(file.path(path.to.02.output, "final_summary_clones_1stDataset_CD45_clusterRes_1.modified.xlsx"))
clone.summary.cd45.modified.raw <- readxl::read_excel(file.path(path.to.02.output, "final_summary_clones_1stDataset_CD45.raw.xlsx")) 

clone.summary.cd45.modified.raw <- clone.summary.cd45.modified.raw %>%
  rowwise() %>%
  mutate(total = CD45exp1_m1 + CD45exp1_m2)

clone.summary.cd45 <- readxl::read_excel(file.path(path.to.02.output, "final_summary_clones_1stDataset_CD45.xlsx"))
```


# Sampling cells and calculate Shannon entropy, GFP samples

## Boxplot: All Shannon entropies

```{r dataset1_all_read_clone_shannon_GFP, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
all.shannon.entropy <- subset(clone.summary.gfp.modified, select = c(CTaa, Shannon.entropy))
all.shannon.entropy %>% ggplot(aes(y = Shannon.entropy)) + geom_boxplot() 

```

## Boxplot: Comparson between real clones and sampled clones Shannon entropies {.tabset}

```{r dataset1_sampling_shannon_GFP, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
all.clone.sizes <- sort(unique(clone.summary.gfp.modified$total))

##### Helper function
calculate_shannon_entropy_sampling_cells <- function(list_of_cells, s.obj){
  N <- length(unique(s.obj$seurat_clusters))
  count_clone_in_cluster <- table(subset(s.obj@meta.data, row.names(s.obj@meta.data) %in% list_of_cells) %>% 
                                    subset(select = c(seurat_clusters))) %>%  as.data.frame()
  count_clone_in_cluster <- count_clone_in_cluster %>% rowwise %>% mutate(Prob = Freq / sum(count_clone_in_cluster$Freq)) %>%
    subset(Prob != 0)
  shannon_entropy <- -sum(count_clone_in_cluster$Prob * log2(count_clone_in_cluster$Prob))/log2(N)
  return(shannon_entropy)
}

for (num.sampling.cells in all.clone.sizes){
  if (file.exists(file.path(path.to.03.output, "validation_GFP_ShannonEntropy", sprintf("validation_Shannon_entropy_sampling_%s.GFP.xlsx", num.sampling.cells))) == FALSE){
    all.cells <- colnames(s.obj.1st.GFP)
    all.sampling.entropies <- c()
      
    for (count in seq(1, 1000)){
      sampling.cells <- sample(all.cells, num.sampling.cells)
      tmp.sampling.shannon.entropy <- calculate_shannon_entropy_sampling_cells(sampling.cells, s.obj.1st.GFP)
      all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
    }
      
    samplingdf <- data.frame(data = all.sampling.entropies)
    colnames(samplingdf) <- c("Shannon.entropy")
    samplingdf$CTaa <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
    samplingdf <- samplingdf[c("CTaa", "Shannon.entropy")]
    samplingdf$case <- "Sampling"
      
    all.shannon.entropy$case <- "real_clone"
      
    validationdf <- rbind(all.shannon.entropy, samplingdf)
    
    writexl::write_xlsx(validationdf, file.path(path.to.03.output, "validation_GFP_ShannonEntropy", sprintf("validation_Shannon_entropy_sampling_%s.GFP.xlsx", num.sampling.cells)))  
  } else {
    validationdf <- readxl::read_xlsx(file.path(path.to.03.output, "validation_GFP_ShannonEntropy", sprintf("validation_Shannon_entropy_sampling_%s.GFP.xlsx", num.sampling.cells)))
    p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
    ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells))
    cat(sprintf("### Sampling %s cells \n", num.sampling.cells))
    print(p)
    cat("\n \n")
  }
}
```


## Double y-axis plot
```{r table_gfp, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
clone.summary.gfp.modified %>% create_dt()
```

### Clones unique in 1 sample


```{r dataset1_double_y_clone_in_unique_1_sample_GFP, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 200
p <- subset(clone.summary.gfp.modified, clone.summary.gfp.modified$Status == "unique_in_1_sample") %>% 
  ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 
p
```

### Clones shared in 2 sample
```{r dataset1_double_y_clone_in_2_samples_GFP, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 200
p <- subset(clone.summary.gfp.modified, clone.summary.gfp.modified$Status == "in_2_samples") %>% 
  ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 

p
```


## New Shannon entropy plot
```{r echo=FALSE, warning=FALSE, include=FALSE, fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj.1st.GFP@meta.data

get_shannon_entropy <- function(x, clone.summary){
  if (nrow( subset(clone.summary, clone.summary$CTaa == x) ) != 0){
    return(subset( clone.summary, clone.summary$CTaa == x )$Shannon.entropy)
  } else {
    return(NA)
  }
}

colnames(clone.summary.gfp.modified.raw) <- c("Clone", "Status", "GFPexp1_m1", "GFPexp1_m2", "total.raw" )
meta.data <- meta.data %>% rowwise() %>% 
  mutate(Shannon.entropy = get_shannon_entropy(CTaa, clone.summary.gfp.modified)) %>% 
  mutate(total = ifelse(is.na(CTaa), NA, subset(clone.summary.gfp.modified.raw, clone.summary.gfp.modified.raw$Clone == CTaa)$total.raw))
  
s.obj.1st.GFP <- AddMetaData(object = s.obj.1st.GFP, metadata = meta.data$Shannon.entropy, col.name = "Shannon.entropy")
s.obj.1st.GFP <- AddMetaData(object = s.obj.1st.GFP, metadata = meta.data$total, col.name = "total")

```

### Clones >= 10 cells, Shannon entropy on UMAP

Clones < 10 cells are displayed in light grey color. 

```{r dataset1_shannon_GFP_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
FeaturePlot(object = s.obj.1st.GFP, reduction = "RNA_UMAP", features = c("Shannon.entropy"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#FF0000", na.value = "lightgray") 
```

### Clones < 10 cells, Clone size on UMAP
Display clone size of all clones < 10 cells on UMAP. 

```{r dataset1_clonesize_lt10_GFP_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj.1st.GFP@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.lt10 = ifelse(CTaa %in% clone.summary.gfp.modified$CTaa == FALSE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj.1st.GFP@meta.data), ]

s.obj.1st.GFP <- AddMetaData(object = s.obj.1st.GFP, metadata = meta.data$total.lt10, col.name = "total.lt10")
FeaturePlot(object = s.obj.1st.GFP, 
            reduction = "RNA_UMAP", features = c("total.lt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

### Clones >= 10 cells, Clone size on UMAP
Display clone size of all clones >= 10 cells on UMAP. 

```{r dataset1_clonesize_gt10_GFP_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
meta.data <- s.obj.1st.GFP@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.gt10 = ifelse(CTaa %in% clone.summary.gfp.modified$CTaa == TRUE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj.1st.GFP@meta.data), ]

s.obj.1st.GFP <- AddMetaData(object = s.obj.1st.GFP, metadata = meta.data$total.gt10, col.name = "total.gt10")
FeaturePlot(object = s.obj.1st.GFP, 
            reduction = "RNA_UMAP", features = c("total.gt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

### All cells. Clone size on UMAP
Display clone size of all clones >= 10 cells on UMAP. 

```{r dataset1_clonesize_all_GFP_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
FeaturePlot(object = s.obj.1st.GFP, 
            reduction = "RNA_UMAP", features = c("total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

### All cells. Clone size on UMAP, log
Display clone size of all clones >= 10 cells on UMAP. 

```{r dataset1_clonesize_all_GFP_samples_logScale, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
s.obj.1st.GFP$log.total <- log10(s.obj.1st.GFP$total)
FeaturePlot(object = s.obj.1st.GFP, 
            reduction = "RNA_UMAP", features = c("log.total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

# Sampling cells and calculate Shannon entropy, CD45 samples 

## Boxplot: All Shannon entropies

```{r dataset1_all_read_clone_shannon_CD45, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
all.shannon.entropy <- subset(clone.summary.cd45.modified, select = c(CTaa, Shannon.entropy))
all.shannon.entropy %>% ggplot(aes(y = Shannon.entropy)) + geom_boxplot() 

```

## Boxplot: Comparson between real clones and sampled clones Shannon entropies {.tabset}

```{r dataset1_sampling_shannon_CD45, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
all.clone.sizes <- sort(unique(clone.summary.cd45.modified$total))

##### Helper function
calculate_shannon_entropy_sampling_cells <- function(list_of_cells, s.obj){
  N <- length(unique(s.obj$seurat_clusters))
  count_clone_in_cluster <- table(subset(s.obj@meta.data, row.names(s.obj@meta.data) %in% list_of_cells) %>% 
                                    subset(select = c(seurat_clusters))) %>%  as.data.frame()
  count_clone_in_cluster <- count_clone_in_cluster %>% rowwise %>% mutate(Prob = Freq / sum(count_clone_in_cluster$Freq)) %>%
    subset(Prob != 0)
  shannon_entropy <- -sum(count_clone_in_cluster$Prob * log2(count_clone_in_cluster$Prob))/log2(N)
  return(shannon_entropy)
}
for (num.sampling.cells in all.clone.sizes){
  if (file.exists(file.path(path.to.03.output, "validation_CD45_ShannonEntropy", sprintf("validation_Shannon_entropy_sampling_%s.CD45.xlsx", num.sampling.cells))) == FALSE){
      all.cells <- colnames(s.obj.1st.CD45)
      all.sampling.entropies <- c()
      
      for (count in seq(1, 1000)){
        sampling.cells <- sample(all.cells, num.sampling.cells)
        tmp.sampling.shannon.entropy <- calculate_shannon_entropy_sampling_cells(sampling.cells, s.obj.1st.CD45)
        all.sampling.entropies <- c(all.sampling.entropies, tmp.sampling.shannon.entropy)
      }
      
      samplingdf <- data.frame(data = all.sampling.entropies)
      colnames(samplingdf) <- c("Shannon.entropy")
      samplingdf$CTaa <- to_vec( for (item in seq(1, 1000)) sprintf("sampling_%s", item))
      samplingdf <- samplingdf[c("CTaa", "Shannon.entropy")]
      samplingdf$case <- "Sampling"
      
      all.shannon.entropy$case <- "real_clone"
      
      validationdf <- rbind(all.shannon.entropy, samplingdf)
      
      writexl::write_xlsx(validationdf, file.path(path.to.03.output, "validation_CD45_ShannonEntropy", sprintf("validation_Shannon_entropy_sampling_%s.CD45.xlsx", num.sampling.cells)))    
  } else {
    validationdf <- readxl::read_xlsx(file.path(path.to.03.output, "validation_CD45_ShannonEntropy", sprintf("validation_Shannon_entropy_sampling_%s.CD45.xlsx", num.sampling.cells)))
    p <- validationdf %>% ggplot(aes(x = case, y = Shannon.entropy)) + geom_boxplot() +
        ggtitle(sprintf("Shannon entropy: Real clone vs sampling %s cells", num.sampling.cells))
    cat(sprintf("### Sampling %s cells \n", num.sampling.cells))
    print(p)
    cat("\n \n")
  }
}
```


## Double y-axis plot
```{r table_cd45, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
clone.summary.cd45.modified %>% create_dt()
```

### Clones unique in 1 sample
```{r dataset1_double_y_clone_in_unique_1_sample_CD45, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 500
p <- subset(clone.summary.cd45.modified, clone.summary.cd45.modified$Status == "unique_in_1_sample") %>% 
  ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  ) 
p
```

### Clones shared in 2 sample
```{r dataset1_double_y_clone_in_2_samples_CD45, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=14, dpi=300, dev='svg'}
coeff <- 500
p <- subset(clone.summary.cd45.modified, clone.summary.cd45.modified$Status == "in_2_samples") %>% 
  ggplot(aes(x = reorder(CTaa, -total))) + 
  geom_point(aes(y = Shannon.entropy)) +  ylim(0, 1) + 
  geom_bar(aes(y = total/coeff), stat = "identity") +  
  theme(axis.text.x = element_blank()) + xlab("Clone") +
  scale_y_continuous(
    name = "Shannon entropy",
    sec.axis = sec_axis(~.*coeff, name="Number of cells") 
  )
p
```

## New Shannon entropy plot

## New Shannon entropy plot
```{r echo=FALSE, warning=FALSE, include=FALSE}
meta.data <- s.obj.1st.CD45@meta.data

get_shannon_entropy <- function(x, clone.summary){
  if (nrow( subset(clone.summary, clone.summary$CTaa == x) ) != 0){
    return(subset( clone.summary, clone.summary$CTaa == x )$Shannon.entropy)
  } else {
    return(NA)
  }
}

colnames(clone.summary.cd45.modified.raw) <- c("Clone", "Status", "CD45exp1_m1", "CD45exp1_m2", "total.raw" )
meta.data <- meta.data %>% rowwise() %>% 
  mutate(Shannon.entropy = get_shannon_entropy(CTaa, clone.summary.cd45.modified)) %>% 
  mutate(total = ifelse(is.na(CTaa), NA, subset(clone.summary.cd45.modified.raw, clone.summary.cd45.modified.raw$Clone == CTaa)$total.raw))

s.obj.1st.CD45 <- AddMetaData(object = s.obj.1st.CD45, metadata = meta.data$Shannon.entropy, col.name = "Shannon.entropy")
s.obj.1st.CD45 <- AddMetaData(object = s.obj.1st.CD45, metadata = meta.data$total, col.name = "total")

```

### Clones >= 10 cells, Shannon entropy on UMAP

Clones < 10 cells are displayed in light grey color. 

```{r dataset1_shannon_CD45_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15, dev='svg'}
FeaturePlot(object = s.obj.1st.CD45, reduction = "RNA_UMAP", features = c("Shannon.entropy"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#FF0000", na.value = "lightgray") + ggtitle("")
```

### Clones < 10 cells, Clone size on UMAP
Display clone size of all clones < 10 cells on UMAP. 

```{r dataset1_clonesize_lt10_CD45_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15, dev='svg'}
meta.data <- s.obj.1st.CD45@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.lt10 = ifelse(CTaa %in% clone.summary.cd45.modified$CTaa == FALSE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj.1st.CD45@meta.data), ]

s.obj.1st.CD45 <- AddMetaData(object = s.obj.1st.CD45, metadata = meta.data$total.lt10, col.name = "total.lt10")
FeaturePlot(object = s.obj.1st.CD45, 
            reduction = "RNA_UMAP", features = c("total.lt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

### Clones >= 10 cells, Clone size on UMAP
Display clone size of all clones >= 10 cells on UMAP. 

```{r dataset1_clonesize_gt10_CD45_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15, dev='svg'}
meta.data <- s.obj.1st.CD45@meta.data %>% rownames_to_column("barcode") %>% subset(select = c(barcode, CTaa, total)) %>%
  rowwise() %>% mutate(total.gt10 = ifelse(CTaa %in% clone.summary.cd45.modified$CTaa == TRUE, total, NA)) %>% column_to_rownames("barcode")

meta.data <- meta.data[row.names(s.obj.1st.CD45@meta.data), ]

s.obj.1st.CD45 <- AddMetaData(object = s.obj.1st.CD45, metadata = meta.data$total.gt10, col.name = "total.gt10")
FeaturePlot(object = s.obj.1st.CD45, 
            reduction = "RNA_UMAP", features = c("total.gt10"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

### All cells, Clone size on UMAP

```{r dataset1_clonesize_all_CD45_samples, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15, dev='svg'}
FeaturePlot(object = s.obj.1st.CD45, 
            reduction = "RNA_UMAP", features = c("total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```


### All cells, Clone size on UMAP, log

```{r dataset1_clonesize_all_CD45_samples_logScale, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15, dev='svg'}
s.obj.1st.CD45$log.total <- log10(s.obj.1st.CD45$total)
FeaturePlot(object = s.obj.1st.CD45, 
            reduction = "RNA_UMAP", features = c("log.total"), pt.size = 1) + scale_color_gradient(low = "#FFFF00", high = "#008b00", na.value = "lightgray") + ggtitle("")
```

# Export final summary table of clones
```{r}
final.clonedf.CD45 <- clone.summary.cd45.modified.raw 
colnames(final.clonedf.CD45) <- c("Clone", "Status", "CD45exp1_m1", "CD45exp1_m2", "total")
final.clonedf.CD45 <- final.clonedf.CD45 %>% rowwise() %>%
  mutate(Shannon.entropy = ifelse( nrow(subset(clone.summary.cd45.modified, clone.summary.cd45.modified$CTaa == Clone)) != 0,
                                   subset(clone.summary.cd45.modified, clone.summary.cd45.modified$CTaa == Clone)$Shannon.entropy, NA)) %>%
  arrange(desc(total))

final.clonedf.GFP <- clone.summary.gfp.modified.raw 
colnames(final.clonedf.GFP) <- c("Clone", "Status", "GFPexp1_m1", "GFPexp1_m2", "total")
final.clonedf.GFP <- final.clonedf.GFP %>% rowwise() %>%
  mutate(Shannon.entropy = ifelse( nrow(subset(clone.summary.gfp.modified, clone.summary.gfp.modified$CTaa == Clone)) != 0,
                                   subset(clone.summary.gfp.modified, clone.summary.gfp.modified$CTaa == Clone)$Shannon.entropy, NA)) %>%
  arrange(desc(total))

writexl::write_xlsx(final.clonedf.GFP, file.path(path.to.save.final.clonedf, "Dataset1_GFP.xlsx"))
writexl::write_xlsx(final.clonedf.CD45, file.path(path.to.save.final.clonedf, "Dataset1_CD45.xlsx"))
```

## GFP
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15}
final.clonedf.GFP %>% create_dt()
```

## CD45
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.height=10, fig.width=15}
final.clonedf.CD45 %>% create_dt()
```


