---
title: "Downstream analysis for the dataset deeper_scRNA-seq_211227"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---
  
`r knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = F)`

`r knitr::opts_chunk$set(fig.width=12, fig.height=8, dpi=500)`


```{r preparation, echo=FALSE, results='hide', include=FALSE, warning=FALSE, message=FALSE}
# _____INPUT THE PREPROCESSED SEURAT OBJECT FROM THE PIPELINE__________
# gc()
# rm(list = ls())

set.seed(42)

analysis.round <- "2nd"

path.to.pipeline.src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline/scRNA_GEX_pipeline"

source(file.path(path.to.pipeline.src, "processes_src", "import_libraries.R"))
source(file.path(path.to.pipeline.src, "processes_src", "helper_functions.R"))
source(file.path(path.to.pipeline.src, "processes_src", "s9_findMarkers.R"))

sample <- sprintf("GFPexp1_m2_%s_round", analysis.round)

outdir <- "/media/hieunguyen/HNSD_MBPro/CRC1382/outdir/LKopplin_OFFICIAL"
input.datadir <- file.path(outdir, "/1stExp_Kopplin/2nd_round")
path.to.VDJ.output <- file.path(outdir, "/1stExp_Kopplin/VDJ_output")

path.to.input <- file.path(input.datadir, sample, "s8a_output")
path.to.output <- file.path(input.datadir, sample, "s9_output")

PROJECT <- "1stExp_Kopplin"

s.obj <- readRDS(file.path(path.to.input, sprintf("%s.output.s8a.rds", PROJECT)))
# s.obj <- subset(s.obj, percent.mt < 0.25)

# _____INPUT THE CLONAL TYPE DATA FROM VDJ DATA ANALYSIS__________

path.to.anno.contigs <- file.path(path.to.VDJ.output, sprintf("annotated_contigs_clonaltype_%s.csv", str_replace(sample, sprintf("_%s_round", analysis.round), "")))

path.to.count.clonaltype <- file.path(path.to.VDJ.output, sprintf("count_clonaltype_%s.csv", str_replace(sample, sprintf("_%s_round", analysis.round), "")))

anno.contigs <- read.csv(path.to.anno.contigs)

count.clonaltype <- read.csv(path.to.count.clonaltype)

anno.contigs <- anno.contigs %>% 
  rowwise %>% 
  mutate(barcode = sprintf("%s_%s", sample, tail(unlist(str_split(barcode, pattern = "_")), 1)))



```
# Preprocessing and QC
## Raw data Quality control  {.tabset}
```{r, results = 'asis', echo = FALSE}
for (plot.name in names(s.obj@misc$all.QC)){
  
  cat('###',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$all.QC[plot.name])
  cat(' \n \n')
  
}

```

Number of cells in each sample after filtering out cells without associated AA sequences: 
  
```{r}
print(table(s.obj@meta.data$name))
```


## Ambient RNA background correction 

### DecontX clusters {.tabset}
```{r, echo = FALSE, message=FALSE, warning=FALSE, results='asis'}

for (plot.name in names(s.obj@misc$ambient.cluster.RNA.plot)){
  
  cat('####',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$ambient.cluster.RNA.plot[plot.name])
  cat(' \n \n')
}
```

### Contamination level in each sample

```{r}
ggplot(s.obj@meta.data, aes(x=name, y=AmbientRNA)) + 
  geom_boxplot()

```


### Ambient RNA contamination level {.tabset}

```{r, results = 'asis', echo = FALSE}
for (plot.name in names(s.obj@misc$ambient.contamination.plot)){
  
  cat('####',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$ambient.contamination.plot[plot.name])
  cat(' \n \n')
}
```


## Descriptive statistics and filtering threshold {.tabset}
This section is devoted to the descriptive statistics of the following varialbes: `nFeature_RNA, nCount_RNA, percent.mt, percent.ribo`. 

```{r, results = 'asis', echo = FALSE}
for (plot.item in c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo")){
  boxplot <- s.obj@meta.data %>% ggplot(aes_string(x = "name", y = plot.item)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16,
                 outlier.size=2, notch=FALSE) +
    ggtitle(sprintf("Boxplot: Distribution of %s in each dataset", plot.item))
  cat('###', plot.item,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  
  print(boxplot)
  
  cat(' \n \n')
}
```

# Clusterings and cluster markers identification
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE)
```

## Clusters

## Cluster markers {.tabset}
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
dir.create(path.to.output, showWarnings = FALSE)


if (file.exists(file.path(path.to.output, "s9_output" ,sprintf("%s.clusterMarkers.s9.rds", PROJECT)))  == FALSE){
  print("Running process S9: Running find cluster markers ...")
  
  cluster.markers <- s9.findClusterMarkers(s.obj, 
                                           path.to.output, 
                                           TRUE,
                                           PROJECT, 
                                           "wilcox")
  
  
} else {
  cluster.markers <- readRDS(file.path(path.to.output, "s9_output", sprintf("%s.clusterMarkers.s9.rds", PROJECT)))
}

diff.cluster.markers <- cluster.markers$diff.markers

```

### Cluster 0
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 0 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```

### Cluster 1
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 1 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```

### Cluster 2
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 2 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```

### Cluster 3
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 3 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```


### Cluster 4
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 4 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```

### Cluster 5
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 5 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```

### Cluster 6
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 6 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```

### Cluster 7
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
subset(diff.cluster.markers, diff.cluster.markers$cluster == 7 & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  subset(avg_log2FC > 0) %>%
  arrange(desc(avg_log2FC)) %>%
  create_dt()
```


## Descriptive statistics + UMAP {.tabset}

### % Mitochondrial
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "percent.mt", label = TRUE, label.color = "red", label.size = 6)
```

### % Ribosome
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "percent.ribo", label = TRUE, label.color = "red", label.size = 6)
```

### % nCount RNA
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "nCount_RNA", label = TRUE, label.color = "red", label.size = 6)
```

### % nGenes 
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "nFeature_RNA", label = TRUE, label.color = "red", label.size = 6)
```


# Clonal overlay

```{r}
top20 <-  head(count.clonaltype, 20)
do.call(patchwork::wrap_plots, lapply(top20$CTaa, function(x) {
  chosen.ctaa <- x
  cell.names <- row.names(subset(slot(s.obj, "meta.data"),
                                 slot(s.obj, "meta.data")$CTaa == chosen.ctaa))
  p <- DimPlot(object = s.obj,
               cells.highlight = cell.names,
               cols.highlight = c("#f77281"),
               cols = "gray", order = TRUE,
               sizes.highlight = 1.5,
               label = TRUE,
               label.box = TRUE) +
    theme(plot.title = element_text(size = 8, face = "bold"), legend.position = "none") +
    ggtitle(chosen.ctaa) 
  return(p)
})) -> subplot_all_top_20_clonaltypes

save_plot_dir <- "/home/hieunguyen/CRC1382/outdir/1stExp_Kopplin/top20_clonaltypes_overlayUMAP"
dir.create(save_plot_dir, showWarnings = FALSE)
dir.create(file.path(save_plot_dir, sample), showWarnings = FALSE)

ggsave(plot = subplot_all_top_20_clonaltypes, filename = file.path(save_plot_dir, sample, sprintf("%s_top20_clonaltypes_overlay_UMAP.png", sample)), device='png', dpi=400, width = 20, height = 15, units = "in")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# remove.clusters <- NULL
# remove.barcodes <- colnames(subset(s.obj, seurat_clusters %in% remove.clusters ))

remove.barcodes <- NULL
dir.create(file.path(path.to.output, "remove_barcodes"), showWarnings = FALSE)

saveRDS(remove.barcodes, file.path(path.to.output, "remove_barcodes", sprintf("%s_remove_barcodes.rds", sample)))
```